\section{Introduction}
Existing introduction is good but needs to be extended a bit with a good motivation as to why we need a model.

\subsection{Previous findings}

Discuss background.
Relationship between TTFs and execution times.
Here we could put Bobby's discussion on neuroticism as a modulating factor.

\section{Related Work}

Take existing related work section.

\section{System Model/Background on WCA}

Explain concepts used in the modeling (execution time, TTF, etc).

\section{Model}

\subsection{Data}

\subsection{Modeling and Implementation}

Discuss some implementation details, present some code.

\subsection{Verification}

Same verification section as before.

\subsection{Task durations}

Compare w.r.t. first-order model.
Explore effects of different levels of neuroticism.

Here we need more content as we are not considering the sampling framework in this paper. Maybe run simulations with different network behaviors/conditions?

Idea: we could use a network simulator to obtain realistic network results. Or even simpler, use local sockets and Linux traffic control utilities to add realistic delays and jitter to the network link.

\subsection{Synthetic Traces}

I think this paper needs to bring back the synthetic traces to make sense.

\section{Validation}

\section{Numerical results}

EdgeDroid 2.0 vs 1.0

Potentially not need the testbed - local simulations.


\section{Discussion}

Optional

\subsection{Neuroticism as a modulating factor}

Maybe this is a better place to put Bobby's discussion.

\subsection{Synthetic traces}

\section{Conclusion}



